---
output: html_document
---

## Salamander Occupancy - Expert Elicitation Manuscript
RMarkdown File

Created by: Rachel Katz (rachelakatz@gmail.com)
Co-authors: Dan Hocking, Evan Grant, Ben Letcher

Created: Feb 01 2016

Last Updated: March 18 2016

#### methods #####
```{r install libraries, echo=FALSE, results='hide'}
library(knitr)
library(xtable)
library(dplyr)
library(tidyr)
library(stringr)
library(RPostgreSQL)
library(DBI)
library(ggplot2)
#library(ggthemes)
library(gridExtra)
library(lme4)
```
```{r connect to db, echo=FALSE, results='hide', eval=FALSE}
# load profile locally to play with packrat
source("~/.Rprofile")

# check user name and password
R.home()
print(options('SHEDS_USERNAME'))
print(options('SHEDS_PASSWORD'))

# set connection to database
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname='sheds_new', host='osensei.cns.umass.edu', user=getOption('SHEDS_USERNAME'), password=getOption('SHEDS_PASSWORD'))
db <- src_postgres(dbname='sheds_new', host='osensei.cns.umass.edu', port='5432', user=options('SHEDS_USERNAME'), password=options('SHEDS_PASSWORD'))
```

```{r download covariates from db, echo=FALSE, results='hide', eval=FALSE}

# column names: featureid |  variable  | value | zone  | riparian_distance_ft 
# zone = upstream or local
# riparian_distance_ft = 200 or NA
# variable =
# featureid = 
# value = 

# get featureids for the deerfield from jeff
tbl_huc12 <- tbl(db, 'catchment_huc12')
df_huc <- tbl_huc12 %>%
     select(huc12, featureid) %>%
     dplyr::mutate(HUC4=substr(huc12, as.integer(1), as.integer(4)),
                   HUC8=substr(huc12, as.integer(1), as.integer(8)),
                   HUC10=substr(huc12, as.integer(1), as.integer(10))) %>%
     dplyr::filter(substr(huc12, as.integer(1), as.integer(8)) == '01080203')
df_huc <- dplyr::collect(df_huc)
head(df_huc)
plot(df_huc$HUC4)  # 1 basin (CT)
plot(df_huc$HUC10) # 5 subwatersheds
plot(df_huc$featureid)
featureids<-c(df_huc$featureid)

# get the covariates for the featureids
tbl_covariates <- tbl(db, 'covariates')
tbl_covariates <- dplyr::filter(tbl_covariates, featureid %in% featureids ) 
df_cov_long<-dplyr::collect(tbl_covariates) # this takes ~ 201pm-
df_covariates<-tidyr::spread(df_cov_long,variable,value)
summary(df_covariates)

df_covariates_m<-as.matrix((df_covariates))
df_covariates_m
write.csv(df_covariates_m,file="df_covariates.csv")
```
```{r import covariate data, echo=FALSE, results='hide'}
# import saved covariate data from db
deerfield.cov.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/df_covariates.csv",fill=TRUE,header=TRUE)
deerfield.cov.data[is.na(deerfield.cov.data)] <- -999

covariates<-deerfield.cov.data %>%
     select(featureid, AreaSqKM,zone,riparian_distance_ft,ann_tmax_c,forest)
str(covariates)
#plot(covariates)

cov.temp<-subset(covariates,ann_tmax_c > -999 & zone=="local")
str(cov.temp) #1088 catchments
plot(cov.temp$ann_tmax_c)
plot(cov.temp$forest)

cov.area<-subset(covariates, ann_tmax_c > -999 & zone=="upstream")
str(cov.area) #1088 catchments
plot(cov.area$AreaSqKM)

cov.ripa<-subset(covariates, riparian_distance_ft == 200 & zone=="local")
str(cov.ripa) #1088 catchments
plot(cov.ripa$forest)

cov<-data.frame(cov.area$AreaSqKM,cov.temp$ann_tmax_c,cov.temp$ann_tmax_c^2,cov.temp$forest,cov.ripa$forest)

# remove greater than 200km2 drainages
cov<-subset(cov,cov$cov.area.AreaSqKM <= 200)
str(cov)
plot(cov)
```
```{r standarize covariates, echo=FALSE, results='hide', eval=FALSE}

beta.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/out_betas.csv",fill=TRUE,header=TRUE)

beta.data

cov.area.std<-(cov$cov.area.AreaSqKM-beta.data[beta.data$predictor=="drainage","mean.cov"])/beta.data[beta.data$predictor=="drainage","sd.cov"]
cov.temp.std<-(cov$cov.temp.ann_tmax_c-beta.data[beta.data$predictor=="temp","mean.cov"])/beta.data[beta.data$predictor=="temp","sd.cov"]
cov.temp2.std<-(cov$cov.temp.ann_tmax_c^2-beta.data[beta.data$predictor=="temp2","mean.cov"])/beta.data[beta.data$predictor=="temp2","sd.cov"]
cov.upla.std<-(cov$cov.temp.forest-beta.data[beta.data$predictor=="upland","mean.cov"])/beta.data[beta.data$predictor=="upland","sd.cov"]
cov.ripa.std<-(cov$cov.ripa.forest-beta.data[beta.data$predictor=="riparian","mean.cov"])/beta.data[beta.data$predictor=="riparian","sd.cov"]
cov.ripa.upla.std<-( (cov$cov.ripa.forest*cov$cov.temp.forest)-beta.data[beta.data$predictor=="riparian.upland.int","mean.cov"])/beta.data[beta.data$predictor=="riparian.upland.int","sd.cov"]

# create matrix of covariates
cov.std.m<-cbind(log(cov.area.std+1))

cov.std.m<-cbind(cov.temp.std)
cov.std.m<-cbind(cov.temp2.std)
cov.std.m<-cbind(cov.temp.std,cov.temp2.std)

cov.std.m<-cbind(cov.upla.std)
cov.std.m<-cbind(cov.ripa.std)
cov.std.m<-cbind(cov.ripa.std,cov.upla.std,cov.ripa.upla.std)

cov.std.m<-cbind(log(cov.area.std+1),cov.temp.std,cov.temp2.std,cov.ripa.std,cov.upla.std,cov.ripa.upla.std)

cov.std.m<-as.matrix(cov.std.m,ncols=6,nrows=985)
str(cov.std.m)
#,cov.temp.std,cov.temp.std^2,cov.upla.std,cov.ripa.std)

# create a vector of betas for each species
dfus.betas<-c(beta.data$defu.mean[2:6],beta.data$defu.mean[8])
#,beta.data$defu.mean[6],beta.data$defu.mean[5])
ebis.betas<-c(beta.data$eubi.mean[2:6],beta.data$eubi.mean[8])
#,beta.data$eubi.mean[6],beta.data$eubi.mean[5])
gpor.betas<-c(beta.data$gypo.mean[2:6],beta.data$gypo.mean[8])
#,beta.data$gypo.mean[6],beta.data$gypo.mean[5])

# calculate probability of occupancy using logit link function
logit.p.dfus<- beta.data$defu.mean[1] + rowSums(cov.std.m * dfus.betas)  
pr_occ.dfus<-1/(1+exp(-(logit.p.dfus)))
logit.p.ebis<- beta.data$eubi.mean[1] + rowSums(cov.std.m * ebis.betas)   
pr_occ.ebis<-1/(1+exp(-(logit.p.ebis)))
logit.p.gpor<- beta.data$gypo.mean[1] + rowSums(cov.std.m * gpor.betas)   
pr_occ.gpor<-1/(1+exp(-(logit.p.gpor)))

cov$dfus_pr_occ<-pr_occ.dfus
cov$ebis_pr_occ<-pr_occ.ebis
cov$gpor_pr_occ<-pr_occ.gpor

ggplot(cov, aes(cov.area.AreaSqKM,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.temp.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.ripa.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.ripa.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.temp.ann_tmax_c,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

```

```{r import expert elicitation data, echo=FALSE, results='hide'}
#import expert data
qall.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/expert_elicitation_questionaire_qall.csv",header=TRUE)

qall2.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/expert_elicitation_questionaire_qall2.csv",stringsAsFactors = TRUE,header=TRUE)
```

#### results #####
#### Salamander Occupancy Model

*Summary of predictor variables across NE catchments*

We used the median and range (min and max or 0, 10, 20, 30, 40, 50, 60, 70, 80, 90, and 100 quantiles) for each covariate across NE catchments to set the covariate space for parameter elicitation. We used the median values of all covariates as the baseline (intercept) and asked experts effects of covariates in relation to this baseline. 


```{r table1, echo=FALSE, results="asis", eval=FALSE}
# elicitation list of ranges
e.list<-matrix(nrow=10,ncol=11,rep(NA,2*11))
e.list[1,1:11]<-c(quantile(catchment.data.sub1$AreaSqKM,seq(0,1,.10)))
e.list[2,1:11]<-c(quantile(catchment.data.sub1$MEAN,seq(0,1,.10)))
e.list[3:10,1:11]<-c(rep(NA,11))
#e.list
colnames(e.list)<-c(seq(0,100, by=10))
rownames(e.list)<-c(
  "drainage_area_sqkm",
  "upstream_%forest",
  "non-buffer_%forest",
  "buffer30m_%forest",
  "upstream_non-riparian_%forest",
  "upstream_riparian30m_%forest",
  "stream_temp_julymean",
  "stream_temp_annualmax",
  "stream_flow_annualmin",
  "stream_flow_annualmax")
table1<-xtable(e.list)
print(table1, type="html")
```

*Occupancy model*

Response variable: probability of species-specific occupancy at the catchment-level

Catchment: Deliniated from hyresolution flow-lines (cite, Kyle). 

Paragraph about covariates expected to influence salamnader occupancy, with the additional hypothesis related to buffer vs. upland land protection (decision model, personal comm).

Model:
pr.occupancy[species,catchment]<-
intercept (overall average occupancy) + 
drainage.area (stream size) + 
stream.temp (july mean or annual max) + 
stream.flow (precip x drainage area; high or low-flow magnitude) + 
riparian.forest (within 30m of riparian) + 
upland.forest (outside of 30m riparian) + 
riparian.forest x upland.forest (interaction term) + 
fish.presence (brook trout; yes or no) + 
residual variance

*Expert results summary*

```{r plot expert results, echo=FALSE,fig.width=6,fig.height=8}

qall.data<-subset(qall2.data,mean_reaches>=0)
str(qall.data)
plot(qall.data[,"mean_reaches"])

# region
pd <- position_dodge(0.1)
p2.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus")
p2.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor")
p2.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis")

grid.arrange(p2.dfus,p2.gpor,p2.ebis,ncol=3)

# drainage
p3.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("dfus")
p3.gpor<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("gpor")
p3.ebis<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("ebis")
grid.arrange(p3.dfus, p3.gpor, p3.ebis)

# temp
p4.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("dfus")
p4.gpor<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("gpor")
p4.ebis<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("ebis")
grid.arrange(p4.dfus, p4.gpor, p4.ebis)

# flow
p5.10.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("dfus, dry year")
p5.190.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("dfus, wet year")

p5.10.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("gpor, dry year")
p5.190.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("gpor, wet year")

p5.10.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("ebis, dry year")
p5.190.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("ebis, wet year")

grid.arrange(p5.10.dfus,p5.190.dfus,
             p5.10.gpor,p5.190.gpor,
             p5.10.ebis,p5.190.ebis,
             ncol=2)

# riparian and upland forest
p6.0.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 0 upland")
p6.25.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 25 upland")
p6.50.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 50 upland")
p6.75.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 75 upland")
p6.100.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 100 upland")
#grid.arrange(p6.0.dfus,p6.25.dfus,p6.50.dfus,p6.75.dfus,p6.100.dfus)

p6.0.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 0 upland")
p6.25.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 25 upland")
p6.50.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 50 upland")
p6.75.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 75 upland")
p6.100.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 100 upland")
#grid.arrange(p6.0.gpor,p6.25.gpor,p6.50.gpor,p6.75.gpor,p6.100.gpor)

p6.0.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 0 upland")
p6.25.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 25 upland")
p6.50.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 50 upland")
p6.75.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 75 upland")
p6.100.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 100 upland")
#grid.arrange(p6.0.ebis,p6.25.ebis,p6.50.ebis,p6.75.ebis,p6.100.ebis)

grid.arrange(p6.0.dfus,p6.25.dfus,p6.50.dfus,p6.75.dfus,p6.100.dfus,
             p6.0.gpor,p6.25.gpor,p6.50.gpor,p6.75.gpor,p6.100.gpor,
             p6.0.ebis,p6.25.ebis,p6.50.ebis,p6.75.ebis,p6.100.ebis,ncol=5)

# trout presence
pd <- position_dodge(0.1)
p7.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus")
p7.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor")
p7.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis")

grid.arrange(p7.dfus,p7.gpor,p7.ebis,ncol=3)

# salamander presence
pd <- position_dodge(0.2)
p8.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("dfus")
p8.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("gpor")
p8.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=c(confidence_reaches), position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("ebis")
grid.arrange(p8.dfus,p8.gpor,p8.ebis,ncol=3)

# old plots
qall.drainage<- qall.data %>%
  dplyr::filter(question=="3" & expert_ID>="0")
p1<-ggplot(qall.drainage,aes(drainage_area_sqkm,mean_reaches,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species))) + geom_point(size=3) + geom_line() 

qall.temp<- qall.data %>%
  dplyr::filter(question=="4" & expert_ID>="4" & pr_occ>=0)
p2<-ggplot(qall.temp, aes(temp,pr_occ,colour=factor(species),shape=factor(expert_ID))) + geom_point(size=3) + geom_line() 

qall.forest<- qall.data %>%
  dplyr::filter(question=="6"& expert_ID>="4")
p3<-ggplot(qall.forest, aes(interaction(riparian,upland),pr_occ,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species,upland)))  + geom_boxplot()

qall.trout<- qall.data %>%
  dplyr::filter(question=="7"& expert_ID>="4")
p4<-ggplot(qall.trout, aes(species,pr_occ,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species))) + geom_point(size=3)


p1
p2
p3
p4
#grid.arrange(p1, p2, p3,p4)
```

*Calculate betas*

```{r calculate betas, echo=FALSE}

qall.data$logit.p<-log(min(-100,(qall.data$mean_reaches+1)/100/(1-((qall.data$mean_reaches+1)/100))))

qall.dfus<-subset(qall.data,qall.data$species=="dfus" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
qall.gpor<-subset(qall.data,qall.data$species=="gpor" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
qall.ebis<-subset(qall.data,qall.data$species=="ebis" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
  
lm1.defu<-lmer(logit.p ~ log(s.area+1) + s.temp +I(s.temp^2) + s.flow + s.freq + s.ripa + s.upla + s.fish + s.ebis + s.gpor + (1|expert_ID), data = qall.dfus)

lm1.gypo<-lmer(logit.p ~ log(s.drainage+1) + s.temp +I(s.temp^2) + s.riparian*s.upland + trout + (1|expert_ID), data = qall.gypo)

lm1.eubi<-lmer(logit.p ~ log(s.drainage+1) + s.temp +I(s.temp^2) + s.riparian*s.upland + trout + (1|expert_ID), data = qall.eubi)


summary(lm1.defu)
qall.defu$pred <- predict(lm1.defu,qall.defu)
qall.defu$resid<- resid(lm1.defu)

summary(lm1.gypo)
qall.gypo$pred <- predict(lm1.gypo,qall.gypo)
qall.gypo$resid<- resid(lm1.gypo)

summary(lm1.eubi)
qall.eubi$pred <- predict(lm1.eubi,qall.eubi)
qall.eubi$resid<- resid(lm1.eubi)

p1<-ggplot(qall.defu, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('DEFU')
p2<-ggplot(qall.gypo, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('GYPO')
p3<-ggplot(qall.eubi, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('EUBI')
p1
p2
p3
#grid.arrange(p1, p2, p3)

qall.new<-rbind(qall.defu,qall.gypo,qall.eubi)

dodge=position_dodge(width=0.5)  

p4<-ggplot(subset(qall.new,qall.new$question==3 & qall.new$expert_ID==4), aes(drainage,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth() + ylim(0,1) + ylab("prob occupancy")+ xlab("catchment drainage area, km2")
p4.5<-ggplot(subset(qall.new,qall.new$question==3 & qall.new$expert_ID==5), aes(drainage,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth() + ylim(0,1) + ylab("prob occupancy")+ xlab("catchment drainage area, km2")

p5<-ggplot(subset(qall.new,qall.new$question==4 & qall.new$expert_ID==4), aes(temp,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth()+ ylim(0,1) + ylab("prob occupancy")+ xlab("mean summer temp, C")
p5.5<-ggplot(subset(qall.new,qall.new$question==4 & qall.new$expert_ID==5), aes(temp,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth()+ ylim(0,1) + ylab("prob occupancy")+ xlab("mean summer temp, C")

p6<-ggplot(subset(qall.new,qall.new$question==6 & qall.new$expert_ID==4), aes(interaction(riparian,upland),1/(1+exp(-pred)),group=interaction(upland,species),color=factor(species))) + geom_point(position=position_dodge(1.1))+ ylim(0,1) + ylab("prob occupancy")+ xlab("percentriparian.percentupland")
p6.5<-ggplot(subset(qall.new,qall.new$question==6 & qall.new$expert_ID==5), aes(interaction(riparian,upland),1/(1+exp(-pred)),group=interaction(upland,species),color=factor(species))) + geom_point(position=position_dodge(1.1))+ ylim(0,1) + ylab("prob occupancy")+ xlab("percentriparian.percentupland")

p7<-ggplot(subset(qall.new,qall.new$question==7 & qall.new$expert_ID==4), aes(trout,1/(1+exp(-pred)),group=interaction(trout,species),color=factor(species))) + geom_boxplot()+ ylim(0,1) + ylab("prob occupancy")+ xlab("brook trout presence")
p7.5<-ggplot(subset(qall.new,qall.new$question==7 & qall.new$expert_ID==5), aes(trout,1/(1+exp(-pred)),group=interaction(trout,species),color=factor(species))) + geom_boxplot()+ ylim(0,1) + ylab("prob occupancy")+ xlab("brook trout presence")

#p4
#p4.5
#p5
#p5.5
#p6
#p6.5
#p7
#p7.5

grid.arrange(p4, p5, p6, p7)
grid.arrange(p4.5, p5.5, p6.5, p7.5)
```

*Plot and export betas*

```{r export betas, echo=FALSE, results='hide'}
betas<-list()
betas$defu<-summary(lm1.defu)
betas$gypo<-summary(lm1.gypo)
betas$eubi<-summary(lm1.eubi)
betas.coef<-data.frame(defu=print(betas$defu$coefficients[,1:2]),gypo=print(betas$gypo$coefficients[,1:2]),eubi=print(betas$eubi$coefficients[,1:2]))
betas.coef$min<-c(NA,
                  min(qall.data$drainage),
                  min(qall.data$temp),
                  min(qall.data$temp^2),
                  min(qall.data$riparian),
                  min(qall.data$upland),
                  min(qall.data$trout),
                  min(qall.data$riparian*qall.data$upland))
betas.coef$max<-c(NA,
                  max(qall.data$drainage),
                  max(qall.data$temp),
                  max(qall.data$temp^2),
                  max(qall.data$riparian),
                  max(qall.data$upland),
                  max(qall.data$trout),
                  max(qall.data$riparian*qall.data$upland))
betas.coef$mean<-c(NA,
                  mean(qall.data$drainage),
                  mean(qall.data$temp),
                  mean(qall.data$temp^2),
                  mean(qall.data$riparian),
                  mean(qall.data$upland),
                  mean(qall.data$trout),
                  mean(qall.data$riparian*qall.data$upland))
betas.coef$sd<-c(NA,
                  sd(qall.data$drainage),
                  sd(qall.data$temp),
                  sd(qall.data$temp^2),
                  sd(qall.data$riparian),
                  sd(qall.data$upland),
                  sd(qall.data$trout),
                  sd(qall.data$riparian*qall.data$upland))
                  
betas.coef$s.min<-c(NA,
                  min(log(qall.data$s.drainage+1)),
                  min(qall.data$s.temp),
                  min(qall.data$s.temp^2),
                  min(qall.data$s.riparian),
                  min(qall.data$s.upland),
                  min(qall.data$trout),
                  min(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.max<-c(NA,
                  max(log(qall.data$s.drainage+1)),
                  max(qall.data$s.temp),
                  max(qall.data$s.temp^2),
                  max(qall.data$s.riparian),
                  max(qall.data$s.upland),
                  max(qall.data$trout),
                  max(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.mean<-c(NA,
                  mean(log(qall.data$s.drainage+1)),
                  mean(qall.data$s.temp),
                  mean(qall.data$s.temp^2),
                  mean(qall.data$s.riparian),
                  mean(qall.data$s.upland),
                  mean(qall.data$trout),
                  mean(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.sd<-c(NA,
                  sd(log(qall.data$s.drainage+1)),
                  sd(qall.data$s.temp),
                  sd(qall.data$s.temp^2),
                  sd(qall.data$s.riparian),
                  sd(qall.data$s.upland),
                  sd(qall.data$trout),
                  sd(qall.data$s.riparian*qall.data$s.upland))
colnames(betas.coef)<-c("defu.mean","defu.sd","gypo.mean","gypo.sd","eubi.mean","eubi.sd","min.cov","max.cov","mean.cov","sd.cov","min.scov","max.scov","mean.scov","sd.scov")
predictor.name<-c("int","drainage","temp","temp2","riparian","upland","trout","riparian.upland.int")
betas.coef$predictor<-predictor.name
write.csv(betas.coef,file="out_betas.csv",col.names=TRUE,row.names=TRUE)
print(betas.coef)
```

```{r plot betas, echo=FALSE, results='hide', fig.height=8, fig.width=5}

betas.mean<-c(betas.coef[1:8,"defu.mean"],betas.coef[1:8,"gypo.mean"],betas.coef[1:8,"eubi.mean"])
betas.sd<-c(betas.coef[1:8,"defu.sd"],betas.coef[1:8,"gypo.sd"],betas.coef[1:8,"eubi.sd"])
betas.sp<-c(rep("defu",8),rep("gypo",8),rep("eubi",8))
betas.name<-c(rep(predictor.name,3))

all.betas<-data.frame(betas.mean,betas.sd,betas.sp,betas.name)
#all.betas

dodge=position_dodge(width=0.5)  
ggplot(all.betas,aes(y=betas.mean,x=betas.name,shape=factor(betas.sp))) + geom_point(aes(shape=betas.sp),position=dodge,size=2) +
geom_errorbar(aes(y=betas.mean,ymin=betas.mean-betas.sd, ymax=betas.mean+betas.sd),size=.5,position=dodge,width=0) +
coord_flip()    

```

```{r occ.model.sim1, echo=FALSE, results='hide',eval=FALSE}
no.catchments=100
no.species=1
no.simulations=20 # number of experts

# create empty vectors and matrices for storing output
sim1.data<-catchment.data.sub1[1:no.catchments,]
pr.occupancy<-alpha<-matrix(rep(NA,no.catchments*no.species*no.simulations))
dim(alpha)<-dim(pr.occupancy)<-c(no.catchments,no.species,no.simulations)
mean.pr.occupancy<-sd.pr.occupancy<-matrix(rep(NA,no.catchments*no.species),nrow=no.catchments,ncol=no.species)
intercept<-b1<-b2<-rep(NA,no.simulations)

# create parameter uncertainty 
# need to calcuate odd-ratios from expert elicitation
# unsure how to add variance/sd in odd-ratios
intercept[1:no.simulations]<-rnorm(no.simulations,mean=log(0.5/(1-0.5)),sd=0.1) # 0.5 average occupancy
b1[1:no.simulations]<-rnorm(no.simulations,log(10),sd=0.1) # 5x more likely to occur for every 1 percent increase in forest cover
b2[1:no.simulations]<-rnorm(no.simulations,log(0.90),sd=0.1) # 10x less likely to occur for every 1 sqkm increase in drainage area

for (species in 1:no.species){
  for (catchment in 1:no.catchments){
    for (sim in 1:no.simulations){
      alpha[catchment,species,sim]<-
      intercept[sim] + b1[sim]*sim1.data$MEAN[catchment] + b2[sim]*sim1.data$AreaSqKM[catchment]
      #+ regional.effect (provinces) 
      #+ stream.temp (july mean or annual max) 
      #+ stream.flow (precip x drainage area; high or low-flow magnitude) 
      #+ riparian.forest (within 30m of riparian)  
      #+ upland.forest (outside of 30m riparian) 
      #+ riparian.forest x upland.forest (interaction term) 
      #+ fish.presence (brook trout; yes or no) 
      #+ residual variance
    pr.occupancy[catchment,species,sim]<-exp(alpha[catchment,species,sim])/(exp(alpha[catchment,species,sim])+1) 
  } # sim
    mean.pr.occupancy[catchment,species]<-mean(pr.occupancy[catchment,species,1:no.simulations])
    sd.pr.occupancy[catchment,species]<-sd(pr.occupancy[catchment,species,1:no.simulations])
 }
}

#head(alpha)
#head(pr.occupancy)
```

```{r plot.parm.uncertainty,echo=FALSE, results="hide",include=FALSE}
par(mfrow=c(3,2))
hist(exp(intercept)/(exp(intercept)+1),main="average occupancy",breaks=10);abline(v=median(exp(intercept)/(exp(intercept)+1)),col="red");
plot(density(exp(intercept)/(exp(intercept)+1)),main="average occupancy");abline(v=median(exp(intercept)/(exp(intercept)+1)),col="red")

hist(exp(b1), main="odds ratio per 1 percent increase in upstream forest",breaks=10);abline(v=median(exp(b1)),col="red")
plot(density(exp(b1)), main="odds ratio per 1 percent increase in upstream forest");abline(v=median(exp(b1)),col="red")

hist(exp(b2), main="odds ratio per 1 sqkm increase in drainage area",breaks=10);abline(v=median(exp(b2)),col="red")
plot(density(exp(b2)), main="odds ratio per 1 sqkm increase in drainage area");abline(v=median(exp(b2)),col="red")
```

Predict mean and sd in occupancy for each catchment

```{r plot.occ,echo=FALSE, results="hide",include=FALSE}
par(mfrow=c(2,2))
#plot(sim1.data$MEAN,sim1.data$AreaSqKM,xlab="percent upstream forest",ylab="drainage area")
plot(mean.pr.occupancy~sim1.data$MEAN,ylim=c(0,1),xlab="percent upstream forest")
plot(mean.pr.occupancy~sim1.data$AreaSqKM,ylim=c(0,1),xlab="drainage area")
plot(sd.pr.occupancy~sim1.data$MEAN,ylim=c(-2,2),xlab="percent upstream forest")
plot(sd.pr.occupancy~sim1.data$AreaSqKM,ylim=c(-2,2),xlab="drainage area")
```

Loop over alterantive climate scenarios

*Climate model uncertainty*

*Buffer-Upland parameter uncertainty*

*Fish-Sal parameter uncertainty*

#### Appendix A. Survey

* SEE SalamanderElictationSurvey_02222016.docx

#### Forecasting Scenarios 
* Outline comparisons of salamander occupancy from this analysis to others from the literature - how will do experts do?

### Results

*Experts responses*

Descripton of how many experts responded.

Parameter estimates (betas) with low vs. high variance among experts.

Model beliefs (weights) with low vs. high variance among experts.

Comparisons across species (similar or different - how?)

Table of betas (mean and sd or scatter plot of results).

Insights about elicitation (base-rates and use of literature we provided)

*Salamander forecasting*

Map of mean expected occupancy for each species with all uncertainties included under current climate, 2040, and 2080 (whichever scenarios we pick)

Histogram of mean expected occupancy for each species with all uncertainties included under current climate, 2040, and 2080 (whichever scenarios we pick). This will show range in expected values (i.e., how many catchments look really bad?)

*Sensitivity analysis*

### Discussion

* expert elicitation process

* insights for salamander conservation (decision-model in prep)

* future directions

### References




