---
output: html_document
---

## Salamander Occupancy - Expert Elicitation Manuscript
RMarkdown File

Created by: Rachel Katz (rachelakatz@gmail.com)
Co-authors: Dan Hocking, Evan Grant, Ben Letcher

Created: Feb 01 2016

Last Updated: March 18 2016

```{r install libraries, echo=FALSE, results='hide'}
library(knitr)
library(xtable)
library(dplyr)
library(tidyr)
library(stringr)
library(RPostgreSQL)
library(DBI)
library(ggplot2)
#library(ggthemes)
library(gridExtra)
library(lme4)

```
```{r connect to db, echo=FALSE, results='hide', eval=FALSE}
# load profile locally to play with packrat
source("~/.Rprofile")

# check user name and password
R.home()
print(options('SHEDS_USERNAME'))
print(options('SHEDS_PASSWORD'))

# set connection to database
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname='sheds_new', host='osensei.cns.umass.edu', user=getOption('SHEDS_USERNAME'), password=getOption('SHEDS_PASSWORD'))
db <- src_postgres(dbname='sheds_new', host='osensei.cns.umass.edu', port='5432', user=options('SHEDS_USERNAME'), password=options('SHEDS_PASSWORD'))

# column names: featureid |  variable  | value | zone  | riparian_distance_ft 
# zone = upstream or local
# riparian_distance_ft = 200 or NA
# variable =
# featureid = 
# value = 

# get featureids for deerfield only from jeff
tbl_huc12 <- tbl(db, 'catchment_huc12')
df_huc <- tbl_huc12 %>%
     select(huc12, featureid) %>%
     dplyr::mutate(HUC4=substr(huc12, as.integer(1), as.integer(4)),
                   HUC8=substr(huc12, as.integer(1), as.integer(8)),
                   HUC10=substr(huc12, as.integer(1), as.integer(10))) %>%
     dplyr::filter(substr(huc12, as.integer(1), as.integer(8)) == '01080203')
df_huc <- dplyr::collect(df_huc)
head(df_huc)
plot(df_huc$huc12)

featureids<-c(df_huc$featureid)
tbl_covariates <- tbl(db, 'covariates')
tbl_covariates <- dplyr::filter(tbl_covariates, featureid %in% featureids ) 
df_cov_long<-dplyr::collect(tbl_covariates) #546pm-
df_covariates<-tidyr::spread(df_cov_long,variable,value)
summary(df_covariates)

df_covariates_m<-as.matrix((df_covariates))
df_covariates_m
write.csv(df_covariates_m,file="df_covariates.csv")
```
```{r import covariate data, echo=FALSE, results='hide'}
# import saved covariate data from db
deerfield.cov.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/df_covariates.csv",fill=TRUE,header=TRUE)
deerfield.cov.data[is.na(deerfield.cov.data)] <- -999

covariates<-deerfield.cov.data %>%
     select(featureid, AreaSqKM,zone,riparian_distance_ft,ann_tmax_c,forest)
str(covariates)
#plot(covariates)

cov.temp<-subset(covariates,ann_tmax_c > -999 & zone=="local")
str(cov.temp) #1088 catchments
plot(cov.temp$ann_tmax_c)
plot(cov.temp$forest)

cov.area<-subset(covariates, ann_tmax_c > -999 & zone=="upstream")
str(cov.area) #1088 catchments
plot(cov.area$AreaSqKM)

cov.ripa<-subset(covariates, riparian_distance_ft == 200 & zone=="local")
str(cov.ripa) #1088 catchments
plot(cov.ripa$forest)

cov<-data.frame(cov.area$AreaSqKM,cov.temp$ann_tmax_c,cov.temp$ann_tmax_c^2,cov.temp$forest,cov.ripa$forest)

# remove greater than 200km2 drainages
cov<-subset(cov,cov$cov.area.AreaSqKM <= 200)
str(cov)
plot(cov)
```
```{r standarize covariates, echo=FALSE, results='hide', eval=FALSE}

beta.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/out_betas.csv",fill=TRUE,header=TRUE)

beta.data

cov.area.std<-(cov$cov.area.AreaSqKM-beta.data[beta.data$predictor=="drainage","mean.cov"])/beta.data[beta.data$predictor=="drainage","sd.cov"]
cov.temp.std<-(cov$cov.temp.ann_tmax_c-beta.data[beta.data$predictor=="temp","mean.cov"])/beta.data[beta.data$predictor=="temp","sd.cov"]
cov.temp2.std<-(cov$cov.temp.ann_tmax_c^2-beta.data[beta.data$predictor=="temp2","mean.cov"])/beta.data[beta.data$predictor=="temp2","sd.cov"]
cov.upla.std<-(cov$cov.temp.forest-beta.data[beta.data$predictor=="upland","mean.cov"])/beta.data[beta.data$predictor=="upland","sd.cov"]
cov.ripa.std<-(cov$cov.ripa.forest-beta.data[beta.data$predictor=="riparian","mean.cov"])/beta.data[beta.data$predictor=="riparian","sd.cov"]
cov.ripa.upla.std<-( (cov$cov.ripa.forest*cov$cov.temp.forest)-beta.data[beta.data$predictor=="riparian.upland.int","mean.cov"])/beta.data[beta.data$predictor=="riparian.upland.int","sd.cov"]

# create matrix of covariates
cov.std.m<-cbind(log(cov.area.std+1))

cov.std.m<-cbind(cov.temp.std)
cov.std.m<-cbind(cov.temp2.std)
cov.std.m<-cbind(cov.temp.std,cov.temp2.std)

cov.std.m<-cbind(cov.upla.std)
cov.std.m<-cbind(cov.ripa.std)
cov.std.m<-cbind(cov.ripa.std,cov.upla.std,cov.ripa.upla.std)

cov.std.m<-cbind(log(cov.area.std+1),cov.temp.std,cov.temp2.std,cov.ripa.std,cov.upla.std,cov.ripa.upla.std)

cov.std.m<-as.matrix(cov.std.m,ncols=6,nrows=985)
str(cov.std.m)
#,cov.temp.std,cov.temp.std^2,cov.upla.std,cov.ripa.std)

# create a vector of betas for each species
dfus.betas<-c(beta.data$defu.mean[2:6],beta.data$defu.mean[8])
#,beta.data$defu.mean[6],beta.data$defu.mean[5])
ebis.betas<-c(beta.data$eubi.mean[2:6],beta.data$eubi.mean[8])
#,beta.data$eubi.mean[6],beta.data$eubi.mean[5])
gpor.betas<-c(beta.data$gypo.mean[2:6],beta.data$gypo.mean[8])
#,beta.data$gypo.mean[6],beta.data$gypo.mean[5])

# calculate probability of occupancy using logit link function
logit.p.dfus<- beta.data$defu.mean[1] + rowSums(cov.std.m * dfus.betas)  
pr_occ.dfus<-1/(1+exp(-(logit.p.dfus)))
logit.p.ebis<- beta.data$eubi.mean[1] + rowSums(cov.std.m * ebis.betas)   
pr_occ.ebis<-1/(1+exp(-(logit.p.ebis)))
logit.p.gpor<- beta.data$gypo.mean[1] + rowSums(cov.std.m * gpor.betas)   
pr_occ.gpor<-1/(1+exp(-(logit.p.gpor)))

cov$dfus_pr_occ<-pr_occ.dfus
cov$ebis_pr_occ<-pr_occ.ebis
cov$gpor_pr_occ<-pr_occ.gpor

ggplot(cov, aes(cov.area.AreaSqKM,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.temp.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.ripa.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.ripa.forest,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

ggplot(cov, aes(cov.temp.ann_tmax_c,y=value)) +
  geom_point(aes(y=dfus_pr_occ,col="blue")) +
  geom_point(aes(y=ebis_pr_occ,col="red")) +
  geom_point(aes(y=gpor_pr_occ,col="green"))

```
```{r import expert elicitation data, echo=FALSE, results='hide'}
#import expert data
qall.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/expert_elicitation_questionaire_qall.csv",header=TRUE)

qall2.data<-read.csv("/Users/katz-umass/github_folder/Trout_Salamander_Decisions/Expert_Elicitation/salamander_elicitation_model/expert_elicitation_questionaire_qall2.csv",stringsAsFactors = TRUE,header=TRUE)
```

### Outline

Overview: Elicite parameter values and model beliefs from experts for predicting effects of climate, land use, and brook-trout presense on catchment-level occupancy of three stream-salamander species occuring throughout the northeastern US. Use this model to forecast current distributions of salamanders under differnce climate scenarios and compare to exisiting published distributions.

* Step 1: build the catchment-level salamander occupancy model
* Step 2: create expert elicitation survey for model parmaeters and beliefs (means only, use 20+ experts to get variance in mean expected values instead of eliciting confidence) 
* Step 3: identify salamander experts based on snow-ball or publication criteria
* Step 4: provide a list of literature for experts to consult prior to elicitation
* Step 5: compare expert opinions and host webinar/call if large discrepancys suggesting lack of understanding of the questions (repeat elicitation if needed)
* Step 6: forecast salamander occupancy based on expert elicitation
* Step 7: conduct sensititivly analysis for major uncertainties (hold all constant at the mean and vary one parmaeter at a time)

### Introduction

*Why use expert knowledge for prediction in conservation*

* Expert knowledge is widely used in conservation because of the complexity of ecological problems, relative lack of empirical data, and imminent nature of decisions. Must act in the face of limited data. 
* Thus, knowledge typically takes the form of mental models without explicit consideration of biases (i.e., experts are called on to express opinions in a qualitative and non-transparent way). 
* In fields where there is extensive expert knowledge, yet little published data, the use of expert information for ecological models is a cost-effective way of making more confident predictions about the effect of management.

* Use of expert elicitation to consolidate the current state of knowledge is valuable when the status of a species is uncertain, yet there is general recognition that the species is sub- ject to multiple, often synergistic, threats to its persistence in the wild.

*Increasing use of expert-knowledge in conservation*

* Increasing use of expert elicitation methods to help inform conservation decisions (identify important hypothesis and parameter uncertainties to resolve through research or monitoring).
* The use of expert knowledge is growing as a tool for synthesizing diverse sources of information into a reliable representation of the current state of scientific knowledge, including its uncertainties when elicited with the same level of rigour pro- vided in the collection and use of empirical data.
* When empirical measures are required for evaluating alternative actions, but are not available, expert judgment can help develop plausible estimates that most effectively inform decisions (Speirs-Bridge et al. 2010; Martin et al. 2012). Expert opinions are valuable in the management of threatened species, where the need to make urgent decisions leaves little time for the collection of further information (?).
* Expert elicitation can be conducted within a probabilistic framework, which accounts for uncertainty (confidence) of multiple experts. Also used to parameterize models in the absence of data and as Bayesian priors to supplement sparse data and improve confidence in predictions (Yamada et al., 2003; Martin et al., 2005; Denham and Mengersen, 2007; Griffiths et al., 2007; Mac Nally, 2007; O’Neill et al., 2008; Low Choy et al., 2009; O’Leary et al., 2009; Murray et al., 2009; James et al., 2010)
* Some chance that expert opinions alone (w/ no data) may not be as useful? Yet, when lacking data, expert opinion does not necessarily offer an improvement (Cox, 2000; Pearce et al., 2001; Seoane et al., 2005).
* Examples of expert elicitation used to parameterize models that inform management decisions (from Martin et al. 2012 and Charney 2012:):

1. Johnson et al. 2010 – use to parameterize Bayesian network to evaluate population viability of cheetahs (X experts).
2. Runge et al. 2011 – use to identify alternative hypotheses and uncertainties affecting mgt of whooping cranes (X experts).
3. O’Neill et al. 2008 – used to quantify trends and range of effects of climate change on polar bears (10 experts).
4. Martin et al. 2005  and Kuhnert et al. 2005 – used to evaluate effects of livestock grazing intensity on austrialian woodland birds in Bayesian GLM (20 experts) 
5. Smith et al. 2011 – Bayesian networks

*What do we mean by "expert" - who are the experts?*

* An expert is someone who has substantial information on a particular topic that is not widely known by others and who is offered deferred to for their knowledge and interpretation.

* Expert knowledge can be defined as substantive information on a particular topic that is not widely known by others (Martin et al., 2012). 

* Experts are distinguished from non-experts by having training or experience with respect to a topic of interest (Fazey et al., 2006).

*Expert elicitation process*

* Expert-elicitation (in general) is composed of 5 steps:
1. Deciding how information will be used (purpose and motivation)
2. Determining what to elicit (and identifying experts; creating a statistical model)
3. Designing elicitation process
4. Performing elicitation
5. Translating elicited information into quantitative statements that be used in ecological models (to ultimately help make decisions).

* Additional important steps to consider:
1. How to work with multiple experts
2. How to combine multiple judgments (treat as equal or weights, average or ranges)
3. Minimizing bias in elicited information
4. Verifying accuracy of expert information

* Why conduct expert elictation for predicting/forecasting stream salamander occupancy?*

* Limited information about amphibians (salamanders) region-wide and evidence that they are currently in decline (cite) and are threatened by climate change (cite).

* Three common stream salamdaner species occur across the northeastern streams (i.e., relevant for NECSC) and are known to co-occur in streams with fishes (brook trout; cite), an important predictor/competitor affecting salamander populations, and are sensitive to buffer and upland forest cover, stream temperature and discharge (and network structure).

* Conduct expert elicitation for Plethedontide: Desmognathus fuscus; dusky salamander, Gyrinophilus porphyriticus; spring salamander, and Eurycea bislineata; two-lined salamander

* Direct threats?
* Indirect threats -  land use (forest cover; urbanization), disease, fish-interactions, climate change (temperature and flow)

* Outstanding hypotheses relevant to conservation and management (personal communication; decision-model in prep): 

1. riparian vs. upland land protection is better for instream biota (downstream). Evidence that upland land protection can influence downstream habitat and species occurrence, yet conservation efforts are largely focused on riparian mgt first, then upland (we'll explore this in decision paper)
2. fish-interaction is strong or weak (brook trout-salamander interactions)
3. species-specific responses (how much do species vary?)
4. stream-network impacts on species-responses? some watersheds are more vulnerable than others (i.e., network structure, buffer vs. upland effects?)

*Now increasing threat of climate change (via temperature?)

* Call for experts to aid in identifying parameters to be used in both predictive and decision models to evaluate range of outcomes of conservation actions.
* Do predictions from experts vary from publications? (maybe compare in analysis or compare in discussion)
* Species distribution models have to make lots of assumptions, we hope experts are considering these assumption and relaxing them, thus providing a more realistic understanding of drivers of occurrence. (be specific here if we mention why SDM models aren't good enough - for using in next co-occuring paper and decision paper)

* Knowledge of that status and potential declines is needed to identify mangagement actions that can be implemented to reduce population declines in the future (act now to mitigate climate/land use threats)

*Current data on these species*

* Current data on these species is patchy?
* No consistant monitoring method or field-studies across their range
* Thus they have high levels of uncertainty when extrapolating to even huc 10 (lowest level of mgt), state, or regional scales.
* Significant geographical gaps in estimates and trend in occupancy (need to come up with a list of current studies). 
* Need address occupancy, trends, threats and effects of potential land-potection (buffer vs. upland) so that headwater stream mgt actions can be confidently employed to impact local and minimize species declines.
* Also, there are diverse experts opinions on occupancy/threats in various regions.


*Objectives of this paper*

Overall Motivation: To use a structured, expert elicitation methods for species in which empirical data are limited and expert opinions may vary (conflict among experts?), such that effectiveness of alternative mgt strategies may vary.

Elicit parameters that can be used for 1) predicting/forecasting salamander occpuancy under climate change, 2) co-occurance (BT + SAL) models under climate change (Hocking et al. forecasting paper), and 3) evaluate robust land-protection strategies for salamanders and trout across watersheds in NE (Katz et al. optimization paper).

Questions:

1. How much do expert opinions agree? Where is the disagreement?
2. Which uncertainties drive forecasts of salamander occurrence (sensitivity analysis; tornado diagram)?
3. Does expert knowledge agree with with field-observations? (why or why not? do we have any? – see publication list that we’ll be sending to experts for potential comparisons).

### Methods

*Identify Experts*

* Ask NEARMI salamander ecologists to name 20 experts
* Use snowball method within survey - who else do you think is more qualified to answer these questions.

Criteria for selecting experts (self-reporting via survey):

Area of expertises (1=none, 2=some, 3=moderate, 4=lots, 5=expert) for each expert:
1. General salamander ecology
2. Landcover effects
3. Riparian effects
4. Stream temperature effects
6. Fish-predation effects
7. Streamflow (drought and floods) effects
8. Occupancy modeling 
9. Abundance modeling
10. Regional expertise (northeast, mid-atlantic, southern)

The X expert participants, who also co-authored this paper, are active salamander researchers in the regions where these species occur and have knowledge of, and experience with, salamander ecology, conservation practice and modelling. (from koala paper)

WE DID NOT USE 4-STEP Elictation on all questions: JUST ONE b/c too many questions and calcuated confidence based on diversity among experts. Martin et al. 2005 (Eco Apps) only elicited the mean, and calculated the mean and precision using mean values across experts (did not elicit lowest, highest and confidence). Use BBN figure (Example BBN for elicitation of variables, see Ban et al. 2015) to identify important conditional probabilities (i.e., riparian x upland, local x upstream) effects?

[We used a structured, four-step process (Speirs-Bridge
et al., 2010; McBride et al., 2012b) to question the experts (Table 1). This procedure helps improve the accuracy of estimates (Burgman et al., 2011), promotes the pooling of individual knowledge and reduces overconfidence (Speirs- Bridge et al., 2010; McBride et al., 2012b).] - koala paper.

We used three steps: pre-survey, webinar-discussion, and post-survey (took 1 month?) + telephone-follow up for extreme values.

*Step 1: Pre-Survey Email questionaire*

* See Appendix A for pre-survey.

*Step 2: Webinar-discusison*
[taken from koala paper]

* Each expert’s first-round answers from the email questionnaire were displayed anonymously to the group. During the 2-hr webinar, the first round of elicited answers was discussed in detail within the group, with experts sharing knowledge from their respective areas of expertise. All experts were given the opportunity to participate in the discussion of each species and allowed to discuss diverse experiences among regions. The experts agreed that this information was influential in guiding their revisions of their estimates in cases where they did not possess knowledge. 

* At the end of the webinar, experts were asked to reconsider their previous assessments in the light of the group discussions. They were given the opportunity to anonymously revise their first-roundassessments. These revised assessments were used in the final estimation of salamander occupancy and effects of environmental variables. 

*Step 3: Post-elicitation*

* Calcualate mean and sd of each betas (used expert as random-effect).

* Aggregation is not always appropriate for a group of judgments (Morgan & Henrion, 1990; Kriegler et al., 2009).

* Assume normally distributed (calcuate median and mean)
* If bimodel distribution - then potentially two hypotheses (camps), so we can go back and elicit beliefs on the two distributions (as structural uncertainty). 
* Betas/predicted relationships between occ and covariates were showed to experts for final comment/review.

*Step 4: Review findings/predictions*

#### Salamander Occupancy Model

*Summary of predictor variables across NE catchments*

We used the median and range (min and max or 0, 10, 20, 30, 40, 50, 60, 70, 80, 90, and 100 quantiles) for each covariate across NE catchments to set the covariate space for parameter elicitation. We used the median values of all covariates as the baseline (intercept) and asked experts effects of covariates in relation to this baseline. 


```{r table1, echo=FALSE, results="asis", eval=FALSE}
# elicitation list of ranges
e.list<-matrix(nrow=10,ncol=11,rep(NA,2*11))
e.list[1,1:11]<-c(quantile(catchment.data.sub1$AreaSqKM,seq(0,1,.10)))
e.list[2,1:11]<-c(quantile(catchment.data.sub1$MEAN,seq(0,1,.10)))
e.list[3:10,1:11]<-c(rep(NA,11))
#e.list
colnames(e.list)<-c(seq(0,100, by=10))
rownames(e.list)<-c(
  "drainage_area_sqkm",
  "upstream_%forest",
  "non-buffer_%forest",
  "buffer30m_%forest",
  "upstream_non-riparian_%forest",
  "upstream_riparian30m_%forest",
  "stream_temp_julymean",
  "stream_temp_annualmax",
  "stream_flow_annualmin",
  "stream_flow_annualmax")
table1<-xtable(e.list)
print(table1, type="html")
```

*Occupancy model*

Response variable: probability of species-specific occupancy at the catchment-level

Catchment: Deliniated from hyresolution flow-lines (cite, Kyle). 

Paragraph about covariates expected to influence salamnader occupancy, with the additional hypothesis related to buffer vs. upland land protection (decision model, personal comm).

Model:
pr.occupancy[species,catchment]<-
intercept (overall average occupancy) + 
drainage.area (stream size) + 
stream.temp (july mean or annual max) + 
stream.flow (precip x drainage area; high or low-flow magnitude) + 
riparian.forest (within 30m of riparian) + 
upland.forest (outside of 30m riparian) + 
riparian.forest x upland.forest (interaction term) + 
fish.presence (brook trout; yes or no) + 
residual variance

*Expert results summary*

```{r plot expert results, echo=FALSE,fig.width=6,fig.height=8}

qall.data<-subset(qall2.data,mean_reaches>=0)
str(qall.data)
plot(qall.data[,"mean_reaches"])

# region
pd <- position_dodge(0.1)
p2.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus")
p2.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor")
p2.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("2") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis")

grid.arrange(p2.dfus,p2.gpor,p2.ebis,ncol=3)

# drainage
p3.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("dfus")
p3.gpor<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("gpor")
p3.ebis<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("3") & mean_reaches>=0), aes(drainage_area_sqkm,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("ebis")
grid.arrange(p3.dfus, p3.gpor, p3.ebis)

# temp
p4.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("dfus")
p4.gpor<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("gpor")
p4.ebis<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("4") & mean_reaches>=0), aes(summer_temp_C,shape=factor(expert_ID))) + geom_point(size=3,aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches))+ ylim(0, 100) + ggtitle("ebis")
grid.arrange(p4.dfus, p4.gpor, p4.ebis)

# flow
p5.10.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("dfus, dry year")
p5.190.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("dfus, wet year")

p5.10.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("gpor, dry year")
p5.190.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("gpor, wet year")

p5.10.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("5") & mean_reaches>=0 & flow_type==10),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("ebis, dry year")
p5.190.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("5") & mean_reaches>=0 & flow_type==190),aes(flow_year_freq,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches,color=factor(stream_width_m))) + geom_line(aes(y=mean_reaches,group=factor(stream_width_m),color=factor(stream_width_m))) +  ylim(0, 100) +  theme(legend.position="none") + ggtitle("ebis, wet year")

grid.arrange(p5.10.dfus,p5.190.dfus,
             p5.10.gpor,p5.190.gpor,
             p5.10.ebis,p5.190.ebis,
             ncol=2)

# riparian and upland forest
p6.0.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 0 upland")
p6.25.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 25 upland")
p6.50.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 50 upland")
p6.75.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 75 upland")
p6.100.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus, 100 upland")
#grid.arrange(p6.0.dfus,p6.25.dfus,p6.50.dfus,p6.75.dfus,p6.100.dfus)

p6.0.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 0 upland")
p6.25.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 25 upland")
p6.50.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 50 upland")
p6.75.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 75 upland")
p6.100.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor, 100 upland")
#grid.arrange(p6.0.gpor,p6.25.gpor,p6.50.gpor,p6.75.gpor,p6.100.gpor)

p6.0.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==0),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 0 upland")
p6.25.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==25),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 25 upland")
p6.50.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==50),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 50 upland")
p6.75.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==75),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 75 upland")
p6.100.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("6") & mean_reaches>=0 & upland_forest==100),aes(riparian_forest,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_line(aes(y=mean_reaches)) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis, 100 upland")
#grid.arrange(p6.0.ebis,p6.25.ebis,p6.50.ebis,p6.75.ebis,p6.100.ebis)

grid.arrange(p6.0.dfus,p6.25.dfus,p6.50.dfus,p6.75.dfus,p6.100.dfus,
             p6.0.gpor,p6.25.gpor,p6.50.gpor,p6.75.gpor,p6.100.gpor,
             p6.0.ebis,p6.25.ebis,p6.50.ebis,p6.75.ebis,p6.100.ebis,ncol=5)

# trout presence
pd <- position_dodge(0.1)
p7.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("dfus")
p7.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("gpor")
p7.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("7") & mean_reaches>=0),aes(expert_ID,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches)) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") + ggtitle("ebis")

grid.arrange(p7.dfus,p7.gpor,p7.ebis,ncol=3)

# salamander presence
pd <- position_dodge(0.2)
p8.dfus<-ggplot(subset(qall.data,species %in% c("dfus") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("dfus")
p8.gpor<-ggplot(subset(qall.data,species %in% c("gpor") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=.1, position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("gpor")
p8.ebis<-ggplot(subset(qall.data,species %in% c("ebis") & question %in% c("8") & mean_reaches>=0),aes(sal_other,mean_reaches,shape=factor(expert_ID))) + geom_point(aes(y=mean_reaches),position=pd) + geom_errorbar(aes(ymin=lower_reaches, ymax=upper_reaches), width=c(confidence_reaches), position=pd) + ylim(0, 100) + theme(legend.position="none") +  ggtitle("ebis")
grid.arrange(p8.dfus,p8.gpor,p8.ebis,ncol=3)

# old plots
qall.drainage<- qall.data %>%
  dplyr::filter(question=="3" & expert_ID>="0")
p1<-ggplot(qall.drainage,aes(drainage_area_sqkm,mean_reaches,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species))) + geom_point(size=3) + geom_line() 

qall.temp<- qall.data %>%
  dplyr::filter(question=="4" & expert_ID>="4" & pr_occ>=0)
p2<-ggplot(qall.temp, aes(temp,pr_occ,colour=factor(species),shape=factor(expert_ID))) + geom_point(size=3) + geom_line() 

qall.forest<- qall.data %>%
  dplyr::filter(question=="6"& expert_ID>="4")
p3<-ggplot(qall.forest, aes(interaction(riparian,upland),pr_occ,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species,upland)))  + geom_boxplot()

qall.trout<- qall.data %>%
  dplyr::filter(question=="7"& expert_ID>="4")
p4<-ggplot(qall.trout, aes(species,pr_occ,colour=factor(species), shape=factor(expert_ID), group=interaction(expert_ID,species))) + geom_point(size=3)


p1
p2
p3
p4
#grid.arrange(p1, p2, p3,p4)
```

*Calculate betas*

```{r calculate betas, echo=FALSE}

qall.data$logit.p<-log(min(-100,(qall.data$mean_reaches+1)/100/(1-((qall.data$mean_reaches+1)/100))))

qall.dfus<-subset(qall.data,qall.data$species=="dfus" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
qall.gpor<-subset(qall.data,qall.data$species=="gpor" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
qall.ebis<-subset(qall.data,qall.data$species=="ebis" & qall.data$expert_ID>=0 & qall.data$logit.p>=0)
  
lm1.defu<-lmer(logit.p ~ log(s.area+1) + s.temp +I(s.temp^2) + s.flow + s.freq + s.ripa + s.upla + s.fish + s.ebis + s.gpor + (1|expert_ID), data = qall.dfus)

lm1.gypo<-lmer(logit.p ~ log(s.drainage+1) + s.temp +I(s.temp^2) + s.riparian*s.upland + trout + (1|expert_ID), data = qall.gypo)

lm1.eubi<-lmer(logit.p ~ log(s.drainage+1) + s.temp +I(s.temp^2) + s.riparian*s.upland + trout + (1|expert_ID), data = qall.eubi)


summary(lm1.defu)
qall.defu$pred <- predict(lm1.defu,qall.defu)
qall.defu$resid<- resid(lm1.defu)

summary(lm1.gypo)
qall.gypo$pred <- predict(lm1.gypo,qall.gypo)
qall.gypo$resid<- resid(lm1.gypo)

summary(lm1.eubi)
qall.eubi$pred <- predict(lm1.eubi,qall.eubi)
qall.eubi$resid<- resid(lm1.eubi)

p1<-ggplot(qall.defu, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('DEFU')
p2<-ggplot(qall.gypo, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('GYPO')
p3<-ggplot(qall.eubi, aes(pred,logit.p), colour="black") + geom_point() + stat_smooth(method="lm", se=TRUE) + ggtitle('EUBI')
p1
p2
p3
#grid.arrange(p1, p2, p3)

qall.new<-rbind(qall.defu,qall.gypo,qall.eubi)

dodge=position_dodge(width=0.5)  

p4<-ggplot(subset(qall.new,qall.new$question==3 & qall.new$expert_ID==4), aes(drainage,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth() + ylim(0,1) + ylab("prob occupancy")+ xlab("catchment drainage area, km2")
p4.5<-ggplot(subset(qall.new,qall.new$question==3 & qall.new$expert_ID==5), aes(drainage,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth() + ylim(0,1) + ylab("prob occupancy")+ xlab("catchment drainage area, km2")

p5<-ggplot(subset(qall.new,qall.new$question==4 & qall.new$expert_ID==4), aes(temp,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth()+ ylim(0,1) + ylab("prob occupancy")+ xlab("mean summer temp, C")
p5.5<-ggplot(subset(qall.new,qall.new$question==4 & qall.new$expert_ID==5), aes(temp,1/(1+exp(-pred)),group=factor(species),color=factor(species))) + geom_point()+ geom_smooth()+ ylim(0,1) + ylab("prob occupancy")+ xlab("mean summer temp, C")

p6<-ggplot(subset(qall.new,qall.new$question==6 & qall.new$expert_ID==4), aes(interaction(riparian,upland),1/(1+exp(-pred)),group=interaction(upland,species),color=factor(species))) + geom_point(position=position_dodge(1.1))+ ylim(0,1) + ylab("prob occupancy")+ xlab("percentriparian.percentupland")
p6.5<-ggplot(subset(qall.new,qall.new$question==6 & qall.new$expert_ID==5), aes(interaction(riparian,upland),1/(1+exp(-pred)),group=interaction(upland,species),color=factor(species))) + geom_point(position=position_dodge(1.1))+ ylim(0,1) + ylab("prob occupancy")+ xlab("percentriparian.percentupland")

p7<-ggplot(subset(qall.new,qall.new$question==7 & qall.new$expert_ID==4), aes(trout,1/(1+exp(-pred)),group=interaction(trout,species),color=factor(species))) + geom_boxplot()+ ylim(0,1) + ylab("prob occupancy")+ xlab("brook trout presence")
p7.5<-ggplot(subset(qall.new,qall.new$question==7 & qall.new$expert_ID==5), aes(trout,1/(1+exp(-pred)),group=interaction(trout,species),color=factor(species))) + geom_boxplot()+ ylim(0,1) + ylab("prob occupancy")+ xlab("brook trout presence")

#p4
#p4.5
#p5
#p5.5
#p6
#p6.5
#p7
#p7.5

grid.arrange(p4, p5, p6, p7)
grid.arrange(p4.5, p5.5, p6.5, p7.5)
```

*Plot and export betas*

```{r export betas, echo=FALSE, results='hide'}
betas<-list()
betas$defu<-summary(lm1.defu)
betas$gypo<-summary(lm1.gypo)
betas$eubi<-summary(lm1.eubi)
betas.coef<-data.frame(defu=print(betas$defu$coefficients[,1:2]),gypo=print(betas$gypo$coefficients[,1:2]),eubi=print(betas$eubi$coefficients[,1:2]))
betas.coef$min<-c(NA,
                  min(qall.data$drainage),
                  min(qall.data$temp),
                  min(qall.data$temp^2),
                  min(qall.data$riparian),
                  min(qall.data$upland),
                  min(qall.data$trout),
                  min(qall.data$riparian*qall.data$upland))
betas.coef$max<-c(NA,
                  max(qall.data$drainage),
                  max(qall.data$temp),
                  max(qall.data$temp^2),
                  max(qall.data$riparian),
                  max(qall.data$upland),
                  max(qall.data$trout),
                  max(qall.data$riparian*qall.data$upland))
betas.coef$mean<-c(NA,
                  mean(qall.data$drainage),
                  mean(qall.data$temp),
                  mean(qall.data$temp^2),
                  mean(qall.data$riparian),
                  mean(qall.data$upland),
                  mean(qall.data$trout),
                  mean(qall.data$riparian*qall.data$upland))
betas.coef$sd<-c(NA,
                  sd(qall.data$drainage),
                  sd(qall.data$temp),
                  sd(qall.data$temp^2),
                  sd(qall.data$riparian),
                  sd(qall.data$upland),
                  sd(qall.data$trout),
                  sd(qall.data$riparian*qall.data$upland))
                  
betas.coef$s.min<-c(NA,
                  min(log(qall.data$s.drainage+1)),
                  min(qall.data$s.temp),
                  min(qall.data$s.temp^2),
                  min(qall.data$s.riparian),
                  min(qall.data$s.upland),
                  min(qall.data$trout),
                  min(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.max<-c(NA,
                  max(log(qall.data$s.drainage+1)),
                  max(qall.data$s.temp),
                  max(qall.data$s.temp^2),
                  max(qall.data$s.riparian),
                  max(qall.data$s.upland),
                  max(qall.data$trout),
                  max(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.mean<-c(NA,
                  mean(log(qall.data$s.drainage+1)),
                  mean(qall.data$s.temp),
                  mean(qall.data$s.temp^2),
                  mean(qall.data$s.riparian),
                  mean(qall.data$s.upland),
                  mean(qall.data$trout),
                  mean(qall.data$s.riparian*qall.data$s.upland))
betas.coef$s.sd<-c(NA,
                  sd(log(qall.data$s.drainage+1)),
                  sd(qall.data$s.temp),
                  sd(qall.data$s.temp^2),
                  sd(qall.data$s.riparian),
                  sd(qall.data$s.upland),
                  sd(qall.data$trout),
                  sd(qall.data$s.riparian*qall.data$s.upland))
colnames(betas.coef)<-c("defu.mean","defu.sd","gypo.mean","gypo.sd","eubi.mean","eubi.sd","min.cov","max.cov","mean.cov","sd.cov","min.scov","max.scov","mean.scov","sd.scov")
predictor.name<-c("int","drainage","temp","temp2","riparian","upland","trout","riparian.upland.int")
betas.coef$predictor<-predictor.name
write.csv(betas.coef,file="out_betas.csv",col.names=TRUE,row.names=TRUE)
print(betas.coef)
```

```{r plot betas, echo=FALSE, results='hide', fig.height=8, fig.width=5}

betas.mean<-c(betas.coef[1:8,"defu.mean"],betas.coef[1:8,"gypo.mean"],betas.coef[1:8,"eubi.mean"])
betas.sd<-c(betas.coef[1:8,"defu.sd"],betas.coef[1:8,"gypo.sd"],betas.coef[1:8,"eubi.sd"])
betas.sp<-c(rep("defu",8),rep("gypo",8),rep("eubi",8))
betas.name<-c(rep(predictor.name,3))

all.betas<-data.frame(betas.mean,betas.sd,betas.sp,betas.name)
#all.betas

dodge=position_dodge(width=0.5)  
ggplot(all.betas,aes(y=betas.mean,x=betas.name,shape=factor(betas.sp))) + geom_point(aes(shape=betas.sp),position=dodge,size=2) +
geom_errorbar(aes(y=betas.mean,ymin=betas.mean-betas.sd, ymax=betas.mean+betas.sd),size=.5,position=dodge,width=0) +
coord_flip()    

```

```{r occ.model.sim1, echo=FALSE, results='hide',eval=FALSE}
no.catchments=100
no.species=1
no.simulations=20 # number of experts

# create empty vectors and matrices for storing output
sim1.data<-catchment.data.sub1[1:no.catchments,]
pr.occupancy<-alpha<-matrix(rep(NA,no.catchments*no.species*no.simulations))
dim(alpha)<-dim(pr.occupancy)<-c(no.catchments,no.species,no.simulations)
mean.pr.occupancy<-sd.pr.occupancy<-matrix(rep(NA,no.catchments*no.species),nrow=no.catchments,ncol=no.species)
intercept<-b1<-b2<-rep(NA,no.simulations)

# create parameter uncertainty 
# need to calcuate odd-ratios from expert elicitation
# unsure how to add variance/sd in odd-ratios
intercept[1:no.simulations]<-rnorm(no.simulations,mean=log(0.5/(1-0.5)),sd=0.1) # 0.5 average occupancy
b1[1:no.simulations]<-rnorm(no.simulations,log(10),sd=0.1) # 5x more likely to occur for every 1 percent increase in forest cover
b2[1:no.simulations]<-rnorm(no.simulations,log(0.90),sd=0.1) # 10x less likely to occur for every 1 sqkm increase in drainage area

for (species in 1:no.species){
  for (catchment in 1:no.catchments){
    for (sim in 1:no.simulations){
      alpha[catchment,species,sim]<-
      intercept[sim] + b1[sim]*sim1.data$MEAN[catchment] + b2[sim]*sim1.data$AreaSqKM[catchment]
      #+ regional.effect (provinces) 
      #+ stream.temp (july mean or annual max) 
      #+ stream.flow (precip x drainage area; high or low-flow magnitude) 
      #+ riparian.forest (within 30m of riparian)  
      #+ upland.forest (outside of 30m riparian) 
      #+ riparian.forest x upland.forest (interaction term) 
      #+ fish.presence (brook trout; yes or no) 
      #+ residual variance
    pr.occupancy[catchment,species,sim]<-exp(alpha[catchment,species,sim])/(exp(alpha[catchment,species,sim])+1) 
  } # sim
    mean.pr.occupancy[catchment,species]<-mean(pr.occupancy[catchment,species,1:no.simulations])
    sd.pr.occupancy[catchment,species]<-sd(pr.occupancy[catchment,species,1:no.simulations])
 }
}

#head(alpha)
#head(pr.occupancy)
```

```{r plot.parm.uncertainty,echo=FALSE, results="hide",include=FALSE}
par(mfrow=c(3,2))
hist(exp(intercept)/(exp(intercept)+1),main="average occupancy",breaks=10);abline(v=median(exp(intercept)/(exp(intercept)+1)),col="red");
plot(density(exp(intercept)/(exp(intercept)+1)),main="average occupancy");abline(v=median(exp(intercept)/(exp(intercept)+1)),col="red")

hist(exp(b1), main="odds ratio per 1 percent increase in upstream forest",breaks=10);abline(v=median(exp(b1)),col="red")
plot(density(exp(b1)), main="odds ratio per 1 percent increase in upstream forest");abline(v=median(exp(b1)),col="red")

hist(exp(b2), main="odds ratio per 1 sqkm increase in drainage area",breaks=10);abline(v=median(exp(b2)),col="red")
plot(density(exp(b2)), main="odds ratio per 1 sqkm increase in drainage area");abline(v=median(exp(b2)),col="red")
```

Predict mean and sd in occupancy for each catchment

```{r plot.occ,echo=FALSE, results="hide",include=FALSE}
par(mfrow=c(2,2))
#plot(sim1.data$MEAN,sim1.data$AreaSqKM,xlab="percent upstream forest",ylab="drainage area")
plot(mean.pr.occupancy~sim1.data$MEAN,ylim=c(0,1),xlab="percent upstream forest")
plot(mean.pr.occupancy~sim1.data$AreaSqKM,ylim=c(0,1),xlab="drainage area")
plot(sd.pr.occupancy~sim1.data$MEAN,ylim=c(-2,2),xlab="percent upstream forest")
plot(sd.pr.occupancy~sim1.data$AreaSqKM,ylim=c(-2,2),xlab="drainage area")
```

Loop over alterantive climate scenarios

*Climate model uncertainty*

*Buffer-Upland parameter uncertainty*

*Fish-Sal parameter uncertainty*

#### Appendix A. Survey

* SEE SalamanderElictationSurvey_02222016.docx

#### Forecasting Scenarios 
* Outline comparisons of salamander occupancy from this analysis to others from the literature - how will do experts do?

### Results

*Experts responses*

Descripton of how many experts responded.

Parameter estimates (betas) with low vs. high variance among experts.

Model beliefs (weights) with low vs. high variance among experts.

Comparisons across species (similar or different - how?)

Table of betas (mean and sd or scatter plot of results).

Insights about elicitation (base-rates and use of literature we provided)

*Salamander forecasting*

Map of mean expected occupancy for each species with all uncertainties included under current climate, 2040, and 2080 (whichever scenarios we pick)

Histogram of mean expected occupancy for each species with all uncertainties included under current climate, 2040, and 2080 (whichever scenarios we pick). This will show range in expected values (i.e., how many catchments look really bad?)

*Sensitivity analysis*

### Discussion

* expert elicitation process

* insights for salamander conservation (decision-model in prep)

* future directions

### References




